---
title: "Introduction to statuAAR"
output: rmarkdown::html_vignette
bibliography: ../inst/REFERENCES.bib
vignette: >
  %\VignetteIndexEntry{Introduction to statuAAR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include=FALSE}
library(statuAAR)
#library(tibble)
knitr::opts_chunk$set(echo = TRUE)
```

## Qickstart with a general workflow.

Obtain a data frame of the available formulas and measurements and decide on 
one or more formulas later.

```{r get formulas}
knitr::kable(statuAAR::getFormulaDataframe())
```

You may check the corresponding documentation, e.g. for breitinger_bach_1965 [@Bach_1965].

```{r check B & B}
?statuAAR::breitinger_bach_1965
```

Load some data into a data frame [e.g. from @Bach_1965] .

```{r read B & B}
Bach1965 <- statuAAR::Bach1965
knitr::kable(head(Bach1965))
```


Prepare tabled data into a long list (statuaar_data_table) and accept the warning 
about the missing ID for each individual.

```{r prep B & B}
dl.bach1965 <- statuAAR::prep.statuaar.data(Bach1965, d.form = "wide",
                   measures.names = "short", sex = "sex", stats = FALSE)
```

Check the measurements using the function provided for general statistics.

```{r check mesaurements B & B}
statuAAR::measures.statistics(dl.bach1965)
```

Get the stature estimations of the provided measurements. The function returns a 
list with a data frame for each formula. You can address each data frame with the 
short name of the corresponding formula. 

```{r get estimates B & B}
knitr::kable(head(statuAAR::getStature(c('bb65'), dl.bach1965)$bb65))
```

## Basic minimal workflow

Create a data frame with measures for one individual per row. If you do not provide
an individual identifier or a sex assignment,  row count and indet. will be added 
in the subsequent data preparation. We create a minimal data frame with data for 
a Neanderthal Man taken from @Pearson_1899[p. 205].

```{r minimal workflow data entry}
x <- data.frame(Fem1 = 445.2, Hum1 = 312, Rad1 = 240)
knitr::kable(x)
```

The column names correspond to the abbreviations used for the bone measurements, 
so no concordance for the names is required (create.measures.concordance()). The 
data can be transfered the wide table into the long format with only one measure 
per row (prep.statuaar.data()).
Furthermore, we do not opt for a general statistic to check for possible errors 
in the data provided.

```{r minimal worflow data preparation}
dl.Neanderthal <- statuAAR::prep.statuaar.data(x, d.form = "wide", stat = FALSE)
knitr::kable(dl.Neanderthal)
```

We now can calculate the estimated stature in various ways.

1. Just use a specific function and get a data frame with the estimated stature, 
the bone measurements used and the alternatives:

```{r minimal worflow stature pearson_1899}
pearson_1899(dl.Neanderthal)
```

2. You can use the function getStature() and provide a vector of formula short 
names and the prepared data frame to get a list containing the corresponding 
data frames. You can convert the list into a single data frame with an additional 
column for the formula. 

```{r minimal worflow staure from vector}
res.Neanderthal <- getStature(c('bo84','bb65','mn09'), dl.Neanderthal)
knitr::kable(getStatureDataframe(res.Neanderthal))
```

If you want to get all possible stature estimations use "all" instead of the 
vector with specific abbreviations. 

```{r minimal worflow staure for all}
res.Neanderthal <- getStature('all', dl.Neanderthal)
knitr::kable(getStatureDataframe(res.Neanderthal))
```


## References

