---
title: "Introduction to statuAAR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to statuAAR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include=FALSE}
library(statuAAR)
knitr::opts_chunk$set(echo = TRUE)
```

## Get started

Get a df of the available formula and measures.

```{r get formulas}
statuAAR::getFormulaDataframe()
```

Check Breitinger & Bach 1965 documentation.

```{r check B & B}
?statuAAR::breitinger_bach_1965
```

Load some data, e.g. Bach1965, into a data.frame.

```{r read B & B}
Bach1965 <- statuAAR::Bach1965
head(Bach1965)
```


Prepare tabled data into a long list (statuaar_data_table).

```{r prep B & B}
dl.bach1965 <- statuAAR::prep.statuaar.data(Bach1965, d.form = "wide",
                   measures.names = "short", sex = "sex", stats = FALSE)
```

Check the measurements using general statistics.

```{r check mesaurements}
statuAAR::measures.statistics(dl.bach1965)
```

Get the stature estimations of the provided measurements. The function returns a list with a data frame for each formula.

```{r get estimates}
head(statuAAR::getStature(c('bb65'), dl.bach1965)$bb65)
```
